\documentclass{article}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{mathtools}
\usepackage{blindtext}
\usepackage{multicol}
\usepackage{graphicx}
\usepackage{placeins}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{booktabs}
\usepackage{float}
\usepackage{bbm}
\graphicspath{ {./images/} }
\usepackage[bottom=0.5cm, right=1.5cm, left=1.5cm, top=1.5cm]{geometry}

\newtheorem{theorem}{Theorem}
\newtheorem{exercise}{Exercise}
\newtheorem{example}{Example}
\newtheorem{definition}{Definition}[section]

\newcommand{\R}{\ensuremath{\mathbb{R}}}
\newcommand{\N}{\ensuremath{\mathbb{N}}}
\newcommand{\Z}{\ensuremath{\mathbb{Z}}}
\newcommand{\p}{\ensuremath{\mathbb{P}}}
\newcommand{\F}{\ensuremath{\mathcal{F}}}
\newcommand{\1}{\ensuremath{\mathbbm{1}}}

\title{Lecture 8}
\author{Gidon Rosalki}
\date{2024-11-25}


\begin{document}
\maketitle
\section{Random variables}\label{sec:Random variables} % (fold)

\begin{definition}[Random variable]
    Let there be $\left(\Omega, \F, \p\right)$. A \textbf{random variable} is
    the function $X : \Omega \to \R$
\end{definition}

\subsubsection{Symbols}\label{sec:Symbols} % (fold)
\begin{itemize}
    \item We will write the collection of events on $\R$ with $\F_\R$
    \item For every $S \in \F_\R$ we will write $\left\{X \in S\right\} = X^{-1} \left(S\right) = \left\{\omega \in \Omega | X(\omega) \in S\right\} \in \F$
\end{itemize}
% subsubsection Symbols (end)

\subsubsection{Examples}\label{sec:Examples} % (fold)
\begin{example}
    $\Omega [6]^2$ (throwing 2 dice), the random variable of the sum of the dice
    is $X(a, b) = a + b$
\end{example}

\begin{example}
    $\Omega = \left\{H, T\right\}^5$ (tossing a coin 5 times), and counting $H$
    is $X(a_1, \dots, a_5) = | \left\{i \in [5] | a_i = H\right\}|$
\end{example}
% subsubsection Examples (end)

When we want to talk about probabilities, we may omit the set symbols. We will
use commas instead of intersection.
\subsubsection{Examples}\label{sec:Examples} % (fold)
\begin{itemize}
    \item $\p \left(X = 5\right) = \p \left(\left\{\omega \in \Omega | X(\omega) = 5\right\}\right)$
    \item $\p \left(X \leq 5, X > 0\right) = \p \left(X \in (0, 5]\right) = \p \left(\left\{\omega \in \Omega | 0 < X(\omega) \leq 5\right\}\right)$
    \item $\p \left(X \leq 5 | X > 0\right) = \p \left(\left\{\omega \in \Omega | X(\omega) \leq 5\right\} | \left\{\omega \in \Omega | 0 < X(\omega)\right\}\right)$
\end{itemize}

\begin{example}[Worked example]
    $\p \left(X \leq 3\right) = \p \left(\left\{\omega \in [6]^2 | X(\omega) \leq 3\right\}\right) = \p \left(\left\{\left(1, 2\right), \left(2, 1\right), \left(1, 1\right)\right\}\right) = \displaystyle\frac{3}{36} = \displaystyle\frac{1}{12}$
\end{example}
% TODO more examples slide 41

\begin{example}[4.4 Gal fished 3 fish]
    The fish come in 3 sizes. A small fish is 50nis, medium is 100, and large is 150.
    So, $\Omega = \left\{S, M, L\right\}^3$ \\
    \begin{align*}
        \text{Amount of small fish: } X &= | \left\{i \in [3] | \omega_i = S\right\}| \\
        \text{Amount of medium fish: } Y &= | \left\{i \in [3] | \omega_i = M\right\}| \\
        \text{Amount of large fish: }Z &= | \left\{i \in [3] | \omega_i = L\right\}| \\
        \text{The amount of non large fish bought: } W_1 &= X + Y \\
        \text{The price of the fish bought: }W_2 &= 50X + 100Y + 150Z \\
        \text{The number of sizes of fish bought: } W_3 &= \min(X, 1) + \min(Y, 1) + \min(Z, 1) \\
    \end{align*}
\end{example}
% subsubsection Examples (end)

\begin{definition}[4.5]
    Let there be $\left(\Omega, \F, \p\right)$, and $A \in \F$. The \textbf{indicator}
    of $A$ is \[
        \1_A(\omega) = \begin{cases}
            1, &\text{ if }\omega \in A\\
            0, &\text{ if }\omega \notin A\\
        \end{cases}
    \]
\end{definition}

\begin{theorem}[4.6]
    Let there be $\left(\Omega, \F, \p\right)$ and $A, B \in \F$

    \begin{enumerate}
        \item $1 - \1_A = \1_{A^c}$
        \item $\1_{A \cap B} = \1_A \1_B$
        \item $\1_{A \cup B} = \max(\1_A, \1_B)$
        \item $0 \notin S \in \F_\R \implies \left\{X \in S\right\} \cap A = \left\{X \cdot \1_A \in S\right\}$
    \end{enumerate}

    \begin{proof}[Proof]
        \begin{enumerate}
            \item \begin{align*}
                    (1 - \1_A)(\omega) &= 1 \\
                    \Leftrightarrow \1_A(\omega) &= 0 \\
                    \Leftrightarrow \omega \notin A \Leftrightarrow \omega \in A^c \Leftrightarrow \1_{A^c} (\omega) = 1
            \end{align*}
            \item \begin{align*}
                    \1_{A \cap B} (\omega) &= 1 \\
                    \Leftrightarrow \omega \in A \cap B \\
                    \Leftrightarrow \omega \in A &\land \omega \in B \\
                    \Leftrightarrow \1_A(\omega) = 1 &\land \1_B(\omega) = 1 \\
                    \Leftrightarrow \1_A(\omega) \1_B(\omega) = 1 \\
            \end{align*}
            \item Done with tables, can write the algebra
            \item TODO Exercise
        \end{enumerate}
    \end{proof}
\end{theorem}

Consider the envelope problem. \\
Given $n$ letters, and $n$ envelopes, we put every letter into an envelope,
without looking what we put where. What is the probability that exactly $0 \leq r \leq n$
letters reach their destinations? \\
\begin{itemize}
    \item The event space is all the arrangements of $[n]$: $\Omega = S_n$
    \item For every $i \in [n]$ we will write $X_i$ to be the envelope to which
        the letter $i$ was inserted: $X_i(\sigma) = \sigma(i)$
    \item For every $i \in [n]$ we will write $Y_i$ to be the indicator that the
        letter $i$ was correctly sent: $Y_i = \1_{X_i = i}$
    \item So, the number of letters correctly sent is $Y = \displaystyle\sum_{i = 1}^{n}Y_i$
\end{itemize}
% section Random variables (end)

\section{Distribution}\label{sec:Distribution} % (fold)
\begin{definition}[4.9 Distribution function]
    Let $\left(\Omega, \F, \p\right)$. Let there be $X: \Omega \to \R$ be a random variable.
    The function $\p_X: \F_\R \to \R$ defined by $\p_X(S) = \p \left(X \in S\right)$
    is the distribution function of $X$
\end{definition}

\begin{theorem}[4.10]
    A distribution function of a random variable $X$ is the probability function on
    $\left(\R, \F_\R\right)$

    \begin{proof}[Proof ]
        $\p_X(\R) = \p \left(X \in \R\right) = \p \left(\omega \in \Omega | X(\omega) \in \R\right) = \p \left(\Omega\right) = 1$ -
        so we have normality. \\
        Let there be $\left(S_n\right)_{n \in \N}$ disjoint events in $\F_\R$.
        $\p_X(\displaystyle\bigsqcup_{n \in \N} S_n) = \p \left(X^{-1} \left(\bigsqcup_{n \in \N} S_n\right)\right)$.
        Set theory exercises: $X^{-1} \left(\displaystyle\bigcup_{A \in I}^{}A\right) = \displaystyle\bigcup_{A \in I}^{}X^{-1} (A)$,
        $A \cap B = \emptyset \implies X^{-1}(A) \cap X^{-1}(B)= \emptyset$. \\
        So, continuing from above, $\p \left(X^{-1} \left(\displaystyle\bigsqcup_{n \in \N} S_n\right)\right) = \displaystyle\sum_{n \in \N}^{}\p \left(X^{-1} \left(S_n\right)\right)
        = \displaystyle\sum_{n \in \N}^{}\p_X(S_n)$

    \end{proof}
\end{theorem}

\begin{definition}[4.11 Discrete random variable]
    A random variable $X$ is \textbf{discrete} if its distribution is discrete,
    ie $\p_X$ is a discrete probability function.
\end{definition}
We will call the function $p_X: \R \to [0, 1]$, which creates the discrete function $\p_X$,
the point discrete distribution function.

\begin{example}
    Let us suppose throwing 2 dice, and taking the sum. Let $X$ be the random variable
    which returns the sum of 2 fair dice.

    \begin{proof}[Solution]
        $\Omega = [6]^2$ has a uniform distribution, and $X((a, b)) = a + b$. (Consider
        a table of dice rolls and their resultant probabilities)

        $\p_X(4 \leq X \leq 6) = p_X(4) + p_X(5) + p_X(6) = \displaystyle\frac{1}{36} \left(3 + 4 + 5\right) = \displaystyle\frac{12}{36}$
    \end{proof}
\end{example}

\subsubsection{Types of distribution}\label{sec:Types of distribution} % (fold)
For building a probability space $\left(\R, \F, \mathcal{D}\right)$ and $X$ a
random variable on the probability space $\left(\Omega, \F, \p\right)$, we will
say that $X$ is distributed by $\mathcal{D}$ if $\p_X = \mathcal{D}$. We will write this
$X \sim \mathcal{D}$

\begin{definition}[4.15]
    A random variable is distributed by the Bernoulli distribution with a probability
    of $p$ if $p_X(1) = p$ and $p_X(0) = 1 - p$
\end{definition}

\begin{example}[4.17 Alice and the queen]
    Alice and the queen of hearts are taking 4 balls from a jug one after the other,
    in which there are 3 red and 2 white. If they remove more red than white, the
    queen wins. Otherwise, Alice wins. What is the probability that the queen wins?

    \begin{proof}[Solution]
        We will assume that the balls are numbered $1$ to $5$, and we will assume that
        the first 3 are red. $\Omega = S_5$ and is uniform. \\
        The queen wins when: $\sigma(4) = 5 \lor \sigma(5) = 5$ (ie, when one of the
        white balls is in the 5 position). \\
        The probability that the queen wins $= X = \1_{\sigma(4) = 5 \lor \sigma(5) = 5} \sim Ber\left(\displaystyle\frac{2}{5}\right)$
        so $\p_X(1) = \p \left(X = 1\right) = \displaystyle\frac{2}{5}$
    \end{proof}
\end{example}

\begin{definition}[4.18 Uniform distribution]
    The random variable $X$ has \textbf{uniform distribution} on the finite set $S \subseteq \R$
    if $\forall x \in S\ p_X(x) = \displaystyle\frac{1}{|S|}$
\end{definition}

We will write $X \sim Unif(S) \lor X \sim U(S)$. Note that $\forall x \in \R \setminus S\ p_X(x) = 0$
% subsubsection Types of distribution (end)

% section Distribution (end)


\end{document}
