\documentclass{article}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{mathtools}
\usepackage{blindtext}
\usepackage{multicol}
\usepackage{graphicx}
\usepackage{placeins}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{booktabs}
\usepackage{float}
\graphicspath{ {./images/} }
\usepackage[bottom=0.5cm, right=1.5cm, left=1.5cm, top=1.5cm]{geometry}

\newtheorem{theorem}{Theorem}
\newtheorem{definition}{Definition}[section]

\newcommand{\R}{\ensuremath{\mathbb{R}}}
\newcommand{\N}{\ensuremath{\mathbb{N}}}
\newcommand{\Z}{\ensuremath{\mathbb{Z}}}
\newcommand{\p}{\ensuremath{\mathbb{P}}}
\newcommand{\F}{\mathcal{F}}

\title{Tutorial 2}
\author{Gidon Rosalki}
\date{2024-11-05}


\begin{document}
\maketitle
\section{Probability spaces} % (fold)
\label{sec:Probability spaces}
\begin{definition}[Event space]
    The \textbf{event space} is the "world" in which all the options are,
    and is symbolised with $\Omega$. An event is written as $\omega$. The
    function $p:\Omega \to \R^+$ which enables $\displaystyle\sum_{\omega \in \Omega}^{}p(w) = 1$
    is called the \textbf{point probability function}.

    The set $supp(p) = \left\{\omega \in \Omega | p(\omega) > 0\right\}$ is called the
    \textbf{support} of $p$. The subset $A \subseteq \Omega$ is called the event,
    and the collection of all events is written $\F$. The probability function
    is defined as the function $\p:\F \to \R^+$ such that:
    \begin{enumerate}
        \item $\p(\Omega) = 1$
        \item $\p(\emptyset) = 0$
        \item For every series of disjoint events $\left\{A_n\right\}_{n \in \N}$
            it is true that $\p(\cup_{n \in \N} A_n) = \displaystyle\sum_{n \in \N}^{}\p(A_n)$
    \end{enumerate}
\end{definition}

$\p$ is called the probability function on $(\Omega, \F)$. We will additionally say
that $\p$ is supported by $A$ if $\p(A) = 1$. Given a point probability function
$p$ on $\Omega$, we may define a probability function on $\Omega, 2^\Omega$ like
so: $\p_p(A) = \displaystyle\sum_{\omega \in \Omega}^{}p(\omega)$. If for a
probability space $(\Omega, \F, \p)$ there exists a point probability function $p:\Omega \to \R^+$,
such that $\p = \p_p$, then $\p$ will be called the \textbf{uniform probability function},
and the space will be called the \textbf{uniform probability space}.

\subsection{Examples} % (fold)
\label{sub:Examples}
\begin{enumerate}
    \item Let us have the event space $\Omega = \left\{1, 2, 3, 4, 5, 6\right\}$,
        with the point probability function $\forall \omega \in \Omega\ p(\omega) = \displaystyle\frac{1}{6}$.
        This example demonstrates throwing a fair die. We will check if $\p_p$ on $\Omega = [6]$
        as written above, and we will get for example
        \[
            \p_p \left(\left\{5, 6\right\}\right) = \displaystyle\frac{1}{6} + \displaystyle\frac{1}{6} = \displaystyle\frac{2}{6} = \displaystyle\frac{1}{3}
        \]
\end{enumerate}
% subsection Examples (end)

\subsection{Uniform probability spaces}
\label{sub:Uniform probability spaces}
\begin{definition}[]
    Let there be $\Omega \ne \emptyset$ a finite set. It is always possible to define the
    probability function $\p$ defined by $\p(A) = \displaystyle\frac{|A|}{|\Omega|} | A \subseteq \Omega$.
    For the probability space that we will get from this $\left(\Omega, 2^\Omega, \p\right)$,
    we call this a uniform probability space, and we call $\p$ a uniform probability
    function.
\end{definition}

Example: If we place 4 black balls in 6 drawers, where we place the balls in drawers
with equal probability. What is the probability that there are no balls in the final
drawer? \\
\textbf{Incorrect} solution: The balls are equivalent in colour, and therefore there
is no consideration to the order of the balls. Therefore, the event space will be
all the ways to order 4 balls in 6 drawers: $\Omega = \left\{\left(x_1, \dots, x_6\right) | \displaystyle\sum_{i = 1}^{6} x_i = 4 \land \forall i\ x_i \geq 0\right\}$.
The point probability function will be $p(\omega) = \displaystyle\frac{1}{|\Omega|}$. We
are interested in the event: $A = \left\{(x_1, \dots, x_5) | \displaystyle\sum_{i = 1}^{5}x_i = 4 \land \forall i\ x_i \geq 0\right\}$.
We get $\p(A) = \displaystyle\frac{|A|}{|\Omega|} = \displaystyle\frac{\binom{4 + 4}{4}}{\binom{4 + 5}{5}} = \displaystyle\frac{5}{9}$.
Why is this solution incorrect? Let us consider 2 balls, and two drawers. There are
2 options for each drawer to have 1 ball, but this solution only counts it once,
because it assumes that order doesn't matter. \\
\textbf{Correct} solution: $\Omega = \left\{\left(x_1, \dots, x_4\right) | \forall i\ x_i \in [6]\right\} = 6^4$.
Additionally $A = \left\{\left(x_1, \dots, x_4\right) | \forall i\ x_i \in [5]\right\} = 5^4$.
So $\p(A) = \displaystyle\frac{|A|}{|\Omega|} = \displaystyle\frac{5^4}{6^4} \approx 0.482$ \\

Example:
\begin{enumerate}
    \item We randomly choose the order of 2 black balls, and 6 white balls. What is
        the chance that the black balls will be adjacent to each other? \\
    \item From a bag of 2 black balls, and 6 white balls, we remove the balls,
        and place them on them in a line. What is the chance that the 2 black balls
        will be adjacent?
\end{enumerate}
Solution:
\begin{enumerate}
    \item $|\Omega| = \binom{8}{2} = 28 \land |A| = \binom{7}{1} = 7$. So $\displaystyle\frac{7}{28} = \displaystyle\frac{1}{4}$.
    \item $|\Omega| = 8! \land |A| = 7! \cdot 2$. So $\p(A) = \displaystyle\frac{7! \cdot 2}{8!} = \displaystyle\frac{1}{4}$
\end{enumerate}

Example: Given a pack of 52 cards, we take 10 cards from the packet. Calculate
what is the chance that exactly 6 clubs? \\
Solution: We will write that $c$ is the collection of cards in the packet,
so $|c| = 52$. Within them $\displaystyle\frac{52}{4} = 13$ are clubs. We will define the probability
space: $\Omega = \left\{\left(c_1, \dots, c_{10}\right) | \forall 1 \leq i \leq 10\ c_i \in c \land c_i \ne c_j\right\}$.
So $|\Omega| = \displaystyle\frac{52!}{42!}$. $A = \text{Exactly 6 cards in} c_1, \dots, c_10 \text{are clubs}$.
So $\p(A) = \displaystyle\frac{|A|}{|\Omega|}$ and $|A| = \binom{10}{6} \cdot \displaystyle\frac{13!}{7!} \cdot \displaystyle\frac{39!}{35!}
\implies \displaystyle\frac{|A|}{|\Omega|} = \displaystyle\frac{1}{100}$
% subsection Uniform probability spaces (end)

\subsection{Infinite probability spaces}
\label{sec:Infinite probability spaces}
\begin{enumerate}
    \item Let $\Omega = \N$. If we define $\p(A) = \displaystyle\frac{|A \cap \left\{1, 2, 3, 4, 5, 6\right\}|}{6}$,
        then this model is the same as a fair die. This probability function is dependent
        on the set $\left\{1, 2, 3, 4, 5, 6\right\}$, despite being defined
        for all $\N$.
    \item Let $\Omega = \N$. Let us define $\p: 2^\Omega \to \R^+$ as follows:
        $\p(A) = \displaystyle\sum_{n \in A}^{}2^{-n}$ and in the following case:
        $\p(\left\{n\right\}) = 2^{-n}$. We will get that $\forall A \subseteq \Omega\ \p(A) \geq 0$.
        $\p(\Omega) = \displaystyle\sum_{n = 1}^{\infty}2^{-n} = 1$ from the sum of
        a geometric series. Given $\left\{A_i\right\}_{i = 1}^{\infty}$ disjoint
        events, and $A = \sqcup_{i = 1}^{\infty}A_i$, then $\p(A) = \displaystyle\sum_{n \in A}^{}2^{-n} =
        \displaystyle\sum_{i = 1}^{\infty}\displaystyle\sum_{n \in A_i}^{}2^{-n} = \displaystyle\sum_{i = 1}^{\infty}\p(A_i)$
\end{enumerate}
% subsection Infinite probability spaces {end}
% section Probability spaces (end)

\section{Multi stage experiments} % (fold)
\label{sec:Multi stage experiments}
\begin{definition}[]
    Let there be $\Omega_1, \Omega_2$ event spaces. A \textbf{two stage experiment}
    is defined by a point probability function $p$ on $\Omega_1$, and also
    for every $\omega_1 \in \Omega_1$ a point probability function $p_{\omega_1}$ on
    $\Omega_2$ defined as the probability of all event in the second step, after we
    got $\omega_1$ from the first step. For a two step experiment,
    it is suitable to use the probability space $\Omega_i = \Omega_1 \times \Omega_2$
    where the point probability function $q: \Omega \to \R^+$ is defined as
    $q((\omega_1, \omega_2)) = p(\omega_1) \cdot o(\omega_2)$.
\end{definition}
% section Multi stage experiments (end)


\end{document}
